Notes: New gaugings in 3d from three-form

### 2022-07-01 Notes_3dH3.tex
Typos corrected and comments added.

Three-form field-strengths are auxiliary in three dimensions and can be integrated out, giving rise to additional gaugings and contributions to the scalar potential, or equivalently leading to new terms in the embedding tensor (see ref.~\cite{Deger:2014ofa,Eloy:2021fhc} for explicit realisations in half-maximal supergravity). We want to generalize this to maximal supergravity. For a given $\E_{8(8)}$ consistent truncations of gauge group $G_{0}$, we ask the question whether it is possible to turn on additional components in the embedding tensor (corresponding to three-form degrees of freedom). If so, the new consistent truncation has gauging $G\supset G_{0}$. We first need to identify which components of the embedding tensor correspond to these degrees of freedom, and to count the number of singlets under $G_{0}$ within them. There will be at least one singlet, corresponding to the gauging of $G_{0}$ in the initial truncation. Any additional singlet would indicate additional parameters to play with, possibly leading to new vacua.

@@

\paragraph{\boldmath $C_{(6)}^{\alpha}$ and $C_{(4)}$ in $\SL(8)_{\rm IIB}\times\R_{+}$} In the $\E_{8(8)}\rightarrow\SL(7)\times\SL(2)\times\R_{+}$ decomposition
\begin{equation}
  \begin{aligned}
    \bm{248} \longrightarrow & \ (\bm{7},\bm{1})_{12} \oplus (\bm{\bar{7}},\bm{2})_{6} \oplus (\bm{\bar{35}},\bm{1})_{6} \oplus (\bm{21},\bm{2})_{3} \oplus (\bm{1},\bm{1})_{0} \oplus (\bm{1},\bm{3})_{0} \oplus (\bm{48},\bm{1})_{0} \\
    & \oplus (\bm{\bar{21}},\bm{2})_{-3} \oplus (\bm{35},\bm{1})_{-6} \oplus (\bm{7},\bm{2})_{-9} \oplus (\bm{\bar{7}},\bm{1})_{-12}
  \end{aligned} 
\end{equation}
of the $\E_{8(8)}$ coordinates $X^{M}$, $C_{(6)}^{\alpha}$ and $C_{(4)}$ sit in the representations $(\bm{7},\bm{2})_{-9}$ and $(\bm{35},\bm{1})_{-6}$, respectively.\footnote{The internal coordinates $y^{m}$ are in the representation $(\bm{7},\bm{1})_{12}$, and $\bm{\bar{7}}^{\wedge 4}=\bm{35}$ and $\bm{\bar{7}}^{\wedge 6}=\bm{7}$.} They originate from the representations $\bm{8}_{3}\oplus\bm{63}_{0}$ and $\bm{56}_{1}$ in $\E_{8(8)}\rightarrow\SL(9)\rightarrow\SL(8)_{\rm IIB}\times\R_{+}$ decomposition
\begin{equation}
  \bm{248} \longrightarrow \bm{80} \oplus \bm{84} \oplus \bm{\bar{84}} \longrightarrow \left[\bm{8}_{3} \oplus \bm{1}_{0} \oplus \bm{63}_{0} \oplus \bm{\bar{8}}_{-3}\right] \oplus \left[\bm{56}_{1} \oplus \bm{28}_{-2}\right] \oplus \left[\bm{\bar{28}}_{2} \oplus \bm{\bar{56}}_{-1}\right],
\end{equation}
whereas the internal coordinates are in the representation $\bm{\bar{28}}_{2}$.

@@

Can be seen as further reduction of four dimensional $\E_{7(7)}$ solutions on a circle. Under $\E_{8(8)}\rightarrow\E_{7(7)}\times\SL(2)\rightarrow\E_{7(7)}\times\R_{+}$,
\begin{equation}
  \bm{248} \longrightarrow (\bm{1},\bm{3}) \oplus (\bm{56},\bm{2}) \oplus (\bm{133},\bm{3})\longrightarrow \bm{1}_{2} \oplus \bm{56}_{1} \oplus \bm{1}_{0} \oplus \bm{133}_{0} \oplus \bm{56}_{-1} \oplus \bm{1}_{-2}.
\end{equation}
The $\E_{7(7)}$ coordinates are in the $\bm{56}_{1}$, and $C_{(6)}$ in $\bm{56}_{-1}$ (11d $\hat{C}_{\hat\mu\hat\nu\hat\rho}$ leads to $C_{\mu\nu\,m}$ in 3d, $H_{\mu\nu\rho\,m}$ dual to purely internal $H_{(7)}$ of potential $C_{(6)}$). Then,
\begin{equation}
    H_{(7)} \subset \bm{56}_{1} \otimes \bm{56}_{-1} = \bm{1}_{0} \oplus \bm{133}_{0} \oplus \bm{1463}_{0} \oplus \bm{1539}_{0}.
\end{equation}
Under $\E_{8(8)}\rightarrow\E_{7(7)}\times\SL(2)\rightarrow\E_{7(7)}\times\R_{+}$, the $\E_{8(8)}$ embedding tensor decomposes as
\begin{equation}
    \bm{3875} \longrightarrow \bm{133}_{2} \oplus \bm{56}_{1} \oplus \bm{912}_{1} \oplus {\color{mydarkorange}\bm{1}_{0}} \oplus {\color{mydarkorange}\bm{133}_{0}} \oplus {\color{mydarkorange}\bm{1539}_{0}} \oplus \bm{56}_{-1} \oplus \bm{912}_{-1} \oplus \bm{133}_{-2}.
\end{equation}

>>>>

Three-form field-strengths are auxiliary in three dimensions and can be integrated out, giving rise to additional gaugings and contributions to the scalar potential, or equivalently leading to new terms in the embedding tensor (see ref.~\cite{Deger:2014ofa,Eloy:2021fhc} for explicit realisations in half-maximal supergravity). We want to generalize this to maximal supergravity. For a given $\E_{8(8)}$ consistent truncations of gauge group $G_{0}$, we ask the question whether it is possible to turn on additional components in the embedding tensor (corresponding to three-form degrees of freedom). If so, the new consistent truncation has gauging $G\supset G_{0}$. We first need to identify which components of the embedding tensor correspond to these degrees of freedom, and to count the number of singlets under $G_{0}$ within them. There will be at least one singlet, corresponding to the gauging of $G_{0}$ in the initial truncation. (\ce{What if the original truncation uses components of the embedding tensor not related to three-forms?}) Any additional singlet would indicate additional parameters to play with, possibly leading to new vacua.

@@

\paragraph{\boldmath $C_{(6)}^{\alpha}$ and $C_{(4)}$ in $\SL(8)_{\rm IIB}\times\R_{+}$} In the $\E_{8(8)}\rightarrow\SL(7)\times\SL(2)\times\R_{+}$ decomposition
\begin{equation}
  \begin{aligned}
    \bm{248} \longrightarrow & \ (\bm{7},\bm{1})_{12} \oplus (\bm{\bar{7}},\bm{2})_{9} \oplus (\bm{\bar{35}},\bm{1})_{6} \oplus (\bm{21},\bm{2})_{3} \oplus (\bm{1},\bm{1})_{0} \oplus (\bm{1},\bm{3})_{0} \oplus (\bm{48},\bm{1})_{0} \\
    & \oplus (\bm{\bar{21}},\bm{2})_{-3} \oplus (\bm{35},\bm{1})_{-6} \oplus (\bm{7},\bm{2})_{-9} \oplus (\bm{\bar{7}},\bm{1})_{-12}
  \end{aligned} 
\end{equation}
of the $\E_{8(8)}$ coordinates $X^{M}$, $C_{(6)}^{\alpha}$ and $C_{(4)}$ sit in the representations $(\bm{7},\bm{2})_{-9}$ and $(\bm{35},\bm{1})_{-6}$, respectively.\footnote{The internal coordinates $y^{m}$ are in the representation $(\bm{7},\bm{1})_{12}$, and $\bm{\bar{7}}^{\wedge 4}=\bm{35}$ and $\bm{\bar{7}}^{\wedge 6}=\bm{7}$.} They originate from the representations $\bm{8}_{3}\oplus\bm{63}_{0}$ and $\bm{56}_{1}$ in $\E_{8(8)}\rightarrow\SL(9)\rightarrow\SL(8)_{\rm IIB}\times\R_{+}$ decomposition
\begin{equation}
  \bm{248} \longrightarrow \bm{80} \oplus \bm{84} \oplus \bm{\bar{84}} \longrightarrow \left[\bm{8}_{3} \oplus \bm{1}_{0} \oplus \bm{63}_{0} \oplus \bm{\bar{8}}_{-3}\right] \oplus \left[\bm{56}_{1} \oplus \bm{28}_{-2}\right] \oplus \left[\bm{\bar{28}}_{2} \oplus \bm{\bar{56}}_{-1}\right],
\end{equation}
whereas the internal coordinates are in the representation $\bm{\bar{28}}_{2}$.

@@

Can be seen as further reduction of four dimensional $\E_{7(7)}$ solutions on a circle. Under $\E_{8(8)}\rightarrow\E_{7(7)}\times\SL(2)\rightarrow\E_{7(7)}\times\R_{+}$,
\begin{equation}
  \bm{248} \longrightarrow (\bm{1},\bm{3}) \oplus (\bm{56},\bm{2}) \oplus (\bm{133},\bm{3})\longrightarrow \bm{1}_{2} \oplus \bm{56}_{1} \oplus \bm{1}_{0} \oplus \bm{133}_{0} \oplus \bm{56}_{-1} \oplus \bm{1}_{-2}.
\end{equation}
The $\E_{7(7)}$ coordinates are in the $\bm{56}_{1}$, and $C_{(6)}$ in $\bm{56}_{-1}$ (11d $\hat{C}_{\hat\mu\hat\nu\hat\rho}$ leads to $C_{\mu\nu\,m}$ in 3d, $H_{\mu\nu\rho\,m}$ dual to purely internal $H_{(7)}$ of potential $C_{(6)}$). Then,
\begin{equation}
    H_{(7)} \subset \bm{56}_{1} \otimes \bm{56}_{-1} = \bm{1}_{0} \oplus \bm{133}_{0} \oplus \bm{1463}_{0} \oplus \bm{1539}_{0}.
\end{equation}
Under $\E_{8(8)}\rightarrow\E_{7(7)}\times\SL(2)\rightarrow\E_{7(7)}\times\R_{+}$, the $\E_{8(8)}$ embedding tensor decomposes as
\begin{equation}
    \bm{3875} \longrightarrow \bm{133}_{2} \oplus \bm{56}_{1} \oplus \bm{912}_{1} \oplus {\color{mydarkorange}\bm{1}_{0}} \oplus {\color{mydarkorange}\bm{133}_{0}} \oplus {\color{mydarkorange}\bm{1539}_{0}} \oplus \bm{56}_{-1} \oplus \bm{912}_{-1} \oplus \bm{133}_{-2}.
\end{equation}
\ce{A priori the $\bm{912}_{1}$ is the 4d embedding tensor, which should support the 4d gaugings, so we need only one singlet in three-form.}
